---
title: "Survival Analysis; Hypothesis testing and Regression"
author: "Hana Akbarnejad"
date: "4/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(survival)
library(KMsurv)
library(survminer)
```


### Hypothesis Testing: Log-rank Test to compare hazard functions between two groups
```{r}
# creating data frame...
surv_data = 
  tibble (
    time = c(4, 12, 15, 21, 23, 2, 6, 8, 10, 19),
    delta = c(1, 0, 1, 0, 1, 1, 0, 0, 1, 1), #0:censored, 1:event
    group = c(1, 1, 1, 1, 1, 2, 2, 2, 2, 2)
  )

#log rank test
survdiff(Surv(time,delta)~group, data=surv_data)

#plotting
plot(survfit(Surv(time,delta)~group, data = surv_data)) 
ggsurvplot( survfit(Surv(time, delta) ~ group, data = surv_data), conf.int=TRUE)
```

We want to test $H_0: h_1(t) = h_2(t)$ versus $H_1:h_1(t) \neq h_2(t)$.

The output of log rank test (*survdiff()* function) shows that p-value is 0.3 which is smaller than 0.05. This means that we fail to reject $H_0$ and conclude that $h_1(t) = h_2(t)$, i.e. hazard function of two groups are not statstically significantly different from each other.

### Comparing death times of kidney transplant patients in different race/sex groups
```{r}
data("kidtran")
kidtran_data = kidtran %>% 
  mutate(
    gender = as.factor(gender),
    race = as.factor(race)
  )

# levels(kidtran_data$gender)  #1:Male, 2:Female
# levels(kidtran_data$race)    #1:white, 2: black

# Surv(kidtran_data$time, kidtran_data$delta,type='right')

#survfit(Surv(time, delta) ~ gender + race, data = kidtran_data)

ggsurvplot(survfit(Surv(time, delta) ~ gender + race, data = kidtran_data), conf.int=FALSE)

# male model
male_surv = survdiff(Surv(time, delta)~ gender + race, data = kidtran_data, subset = (gender == 1))
male_surv
#female model
female_surv = survdiff(Surv(time, delta)~ gender + race, data = kidtran_data, subset = (gender == 2))
female_surv

#male plot
ggsurvplot(survfit(Surv(time, delta) ~ race+gender, data = kidtran_data, subset = (gender == 1) ), conf.int=FALSE)
#female plot
ggsurvplot(survfit(Surv(time, delta) ~ race+gender, data = kidtran_data, subset = (gender == 2) ), conf.int=FALSE)
```

Comparing survival of male subjects, we can observe that p-value is 0.8 which is greater than 0.05. This means that we fail to reject H_0 and conclude that there is no statistically significant difference in survival of white men versus black men. We can also observe that in the plot that shows survival plot of different races in male subjects, the curve are pretty close to each other.

However, this p-value is 0.03 for female subjects, which means that we should reject null hypothesis and conclude that there is a statistically meaningful difference in survival of white versus black females. this difference is also reflected in survival plot and we can see that survival of black women is significantly lower than white women throught the study.

### Proportional Hazards(PH) Model, survival of different stages of Larynx Cancer patients
```{r}
data("larynx")
larynx_data = larynx %>% 
  mutate(
    z1 = if_else(stage == 2, 1, 0),
    z2 = if_else(stage == 3, 1, 0),
    z3 = if_else(stage == 4, 1, 0),
    z4 = age,
    z1 = as.factor(z1),
    z2 = as.factor(z2),
    z3 = as.factor(z3),
  ) %>% 
  select(time, diagyr, delta, z1:z4)

ph_fir = coxph(Surv(time,delta)~z1+z2+z3+z4+z1*z4,data=larynx_data,ties='breslow')
ph_fir
```

The model is:
$h_i(t)=h_0(t) e^{\beta_1 Z_1 + \beta_2 Z_2 + \beta_3 Z_3 + \beta_4 Z_4 + \beta_5 Z1 *  Z4 }$

$$h_i(t)=h_0(t) e^{-7.38Z_1 + 0.62Z_2 + 1.75Z_3 + 0.006 Z_4 + 0.11Z1 *  Z4 }$$

$\beta_1 + \beta_5:$ The log hazard ratio of time to death or censoring between participants in stageII versus those in stageI is -7.27 (-7.38 + 0.11), for a specific age(?).

$\beta_2:$ The log hazard ratio of time to death or censoring between participants in stageIII versus those in stageI is  0.62, holding age constant.

$\beta_3:$ The log hazard ratio of time to death or censoring between participants in stageIV versus those in stageI is  1.75, holding age constant.

$\beta_4:$

